---
title: "final_project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Basic Part

###2.1 Backward returns

```{r}
prices = read.csv('final_project-1.csv')
head(prices)
```

```{r}
# Convert times from range 0 -> T - 1 to 1 -> T
prices[, 1] = prices[, 1] + 1
summary(prices)
```

```{r}
I = c(1, 2, 3)
H = c(3, 10, 30)

BRs = matrix(nrow=nrow(prices), ncol=length(I)*length(H))
```

```{r}
col_num = 1
col_names = rep(0, length(I)*length(H))

# Calculate backwards returns on each Asset
for(i in I) {
  asset_prices = prices[, i+1]
  
  # h-minute backward returns
  for(h in H) {
    # [s(t) - s(max(t - h), 1)] / s(max(t - h), 1)
    prev_returns = asset_prices[pmax(prices[, 1] - h, 1)]
    returns = round((asset_prices - prev_returns) / prev_returns, 4)
    BRs[, col_num] = returns
    
    col_names[col_num] = paste("Asset_",i,"_BRet_",h, sep="")
    col_num = col_num + 1
  }
}
```

```{r}
# Output to bret.csv
BRet = data.frame(BRs)
colnames(BRet) <- col_names
write.csv(BRet, 'bret.csv', row.names = FALSE)
```

### 2.2 Rolling correlation

```{r}
Rho_m = matrix(nrow=nrow(BRet), ncol=length(I))
```

```{r}
# 3 weeks backward rolling correlation
interval = 21 * 24 * 60

col_num = 1
col_names = rep(0, length(I))

for (i in seq(1,3)) {
  j = (i %% 3) + 1
  
  # Swap vars if i > j
  if (i > j) { i = (j - i) + (j = i) }
  
  asset_i_BRet_3 = BRet[, (3 * i) - 2]
  asset_j_BRet_3 = BRet[, (3 * j) - 2]
  
  rcor <- function(t) {
    lower = max(t - interval, 1)
  
    return(round(cor(asset_i_BRet_3[lower:t], asset_j_BRet_3[lower:t]), 4))
  }
  
  Rho_m[, col_num] <- mapply(rcor, prices[, 1])
  col_names[col_num] = paste("Rho_", i, "_", j, sep="")
  col_num = col_num + 1
}
```

```{r}
Rho <- data.frame(Rho_m)
colnames(Rho) <- col_names
write.csv(Rho, 'corr.csv', row.names = FALSE)
```


### 2.3 Linear Regression

```{r}
h = 10

# Use asset 1 prices
asset_prices = prices[, 2]

# 
future_returns = asset_prices[pmin(prices[, 1] + h, length(asset_prices))]
returns = round((future_returns - asset_prices) / asset_prices, 4)

FRet = data.frame("Asset_1_Fret_10"=returns)

fret_bret <- cbind(FRet, BRet)
```

```{r}
train_proportion = 0.7

# Select first 70% of data as training data
train_size <- floor(nrow(fret_bret) * train_proportion)
train_ids = 1:train_size

fret_train <- fret_bret[train_ids, ]
fret_test <- fret_bret[-train_ids, ]

fret_bret.lm <- lm(Asset_1_Fret_10 ~ ., data=fret_train)
summary(fret_bret.lm)
```

The backward returns for Asset 2 and Asset 3 are significant in predicting the forward return of Asset 1 except for the 30-minute backwards returns for Asset 3.

```{r}
pred_train_lm = predict(fret_bret.lm, fret_train)
pred_test_lm = predict(fret_bret.lm, fret_test)

in_sample = cor(pred_train_lm, fret_train$Asset_1_Fret_10)
out_sample = cor(pred_test_lm, fret_test$Asset_1_Fret_10)

cat("In-sample correlation for regression:", in_sample,
    "\nOut-sample correlation for regression:", out_sample)
```

```{r}
# Plot three-week backward rolling correlation between predicted 
# forward returns and actual forward returns

ignore = library(roll)

pred_lm = predict(fret_bret.lm, fret_bret)
rcors_fb <- roll_cor(fret_bret$Asset_1_Fret_10, pred_lm, interval, min_obs = 1)

plot(rcors_fb, main="Rolling 3-week backward correlation between actual and predicted 10 minute forward returns", xlab="Time t (minutes)", ylab="Correlation")
```

We observe the correlation structure to be approximately stationary over the year with mean correlation at around 0.05 and constant variance.


### 2.4 KNN

```{r}
library(FNN)
k_range = c(5, 25, 125, 625, 1000)

trainMSE = c() 
testMSE = c()

# Test multiple values of K
for (i in 1:length(k_range)) {
    # Training MSE
    knnTrain <- knn.reg(train = fret_train[, 2:10], y = fret_train[, 1],
                        test = fret_train[, 2:10], k = k_range[i])
    trainMSE[i] <- mean((fret_train[, 1] - knnTrain$pred)^2)
    
    # Test MSE
    knnTest <- knn.reg(train = fret_train[, 2:10], y = fret_train[, 1],
                       test = fret_test[, 2:10], k = k_range[i])
    testMSE[i] <- mean((fret_test[, 1] - knnTest$pred)^2)
}
```

```{r}
plot(trainMSE ~ I(1/k_range), type="b", lwd = 2, col = "blue",
     main = "Training and Test MSE for KNN", xlab = "1/K", 
     ylab = "MSE", ylim=c(5e-6, 1e-5))
lines(testMSE ~ I(1/k_range), type = "b", lwd = 2, col = "red")

# Legend
legend("bottomleft", legend =c("Training KNN", "Test KNN"), 
       cex = 0.8, col =c("blue", "red"), lwd =c(2, 2), 
       pch =c(1, 1), lty =c(1, 1))
```

```{r}
best_k = k_range[which.min(testMSE)]

# Generate predictions for the whole year with best K.
# Last iteration in loop was for k = 1000
knn_pred = c(knnTrain, knnTest)
# plot(knn_pred$pred)

knn_in_sample = cor(knnTrain$pred, fret_train$Asset_1_Fret_10)
knn_out_sample = cor(knnTest$pred, fret_test$Asset_1_Fret_10)

cat("In-sample correlation:", knn_in_sample,
    "\nOut-sample correlation:", knn_out_sample)
```


### 2.5 Ridge and LASSO

```{r}
I = c(1, 2, 3)
H = c(3, 10, 30, 60, 120, 180, 240, 360, 480, 600, 720, 960, 1200, 1440)

BRs = matrix(nrow=nrow(prices), ncol=length(I)*length(H))
```

```{r}
col_num = 1
col_names = rep(0, length(I)*length(H))

# Calculate backwards returns on each Asset
for(i in I) {
  asset_prices = prices[, i+1]
  
  # h-minute backward returns
  for(h in H) {
    # [s(t) - s(max(t - h), 1)] / s(max(t - h), 1)
    prev_returns = asset_prices[pmax(prices[, 1] - h, 1)]
    returns = round((asset_prices - prev_returns) / prev_returns, 4)
    BRs[, col_num] = returns
    
    col_names[col_num] = paste("Asset_",i,"_BRet_",h, sep="")
    col_num = col_num + 1
  }
}
```

```{r}
BRet_ridge = data.frame(BRs)
colnames(BRet_ridge) <- col_names

FBret_ridge <- cbind(FRet, BRet_ridge)
```

```{r}
library(glmnet)

# Same training and testing samples as before
FBret_x = model.matrix(Asset_1_Fret_10 ~ ., FBret_ridge)[, -1]
FBret_y = FBret_ridge$Asset_1_Fret_10

grid = 10^seq(2, -6, length=100)
ridge.mod = glmnet(FBret_x[train_ids,], FBret_y[train_ids], alpha=0, lambda=grid)
lasso.mod = glmnet(FBret_x[train_ids,], FBret_y[train_ids], alpha=1, lambda=grid)

testMSEs_ridge = c()
testMSEs_lasso = c()

for (lam in grid) {
  ridge_pred = predict(ridge.mod, s=lam, FBret_x[-train_ids, ])
  testMSEs_ridge = c(testMSEs_ridge, mean((FBret_y[-train_ids] - ridge_pred)^2))
  
  lasso_pred = predict(lasso.mod, s=lam, FBret_x[-train_ids, ])
  testMSEs_lasso = c(testMSEs_lasso, mean((FBret_y[-train_ids] - lasso_pred)^2))
}
```

```{r}
best_lam_ridge = grid[which.min(testMSEs_ridge)]
best_lam_lasso = grid[which.min(testMSEs_lasso)]

best_lam_ridge
best_lam_lasso

# Generate prediction for whole year
ridge_pred = predict(ridge.mod, s=best_lam_ridge, FBret_x)
lasso_pred = predict(lasso.mod, s=best_lam_lasso, FBret_x)
```

```{r}
in_sample_ridge = cor(ridge_pred[train_ids], FBret_y[train_ids])
out_sample_ridge = cor(ridge_pred[-train_ids], FBret_y[-train_ids])

cat("In-sample correlation for ridge:", in_sample_ridge,
    "\nOut-sample correlation for ridge:", out_sample_ridge)
```

```{r}
in_sample_lasso = cor(lasso_pred[train_ids], FBret_y[train_ids])
out_sample_lasso = cor(lasso_pred[-train_ids], FBret_y[-train_ids])

cat("In-sample correlation for lasso:", in_sample_lasso,
    "\nOut-sample correlation for lasso:", out_sample_lasso)
```


### 2.6 PCR

```{r}
library(pls)

FR_pcr <- pcr(Asset_1_Fret_10 ~ ., data = FBret_ridge[train_ids, ], scale = FALSE, validation = "none")
```

```{r}
summary(FR_pcr)
```

```{r}
testMSEs_pcr = c()

for(i in 1:(ncol(FBret_ridge)-1)) {
  FR_pcr.pred <- predict(FR_pcr, FBret_ridge[-train_ids, ], ncomp = i)
  PCRTestMSE <- mean((FR_pcr.pred - FBret_ridge[-train_ids, "Asset_1_Fret_10"])^2)
  testMSEs_pcr = append(testMSEs_pcr, PCRTestMSE)
}
```

```{r}
which.min(testMSEs_pcr)

# Generate predictions for the whole year
FR_pcr.pred <- predict(FR_pcr, FBret_ridge, ncomp = which.min(testMSEs_pcr))

in_sample_pcr = cor(FR_pcr.pred[train_ids], FBret_y[train_ids])
out_sample_pcr = cor(FR_pcr.pred[-train_ids], FBret_y[-train_ids])

cat("In-sample correlation for PCR:", in_sample_pcr,
    "\nOut-sample correlation for PCR:", out_sample_pcr)
```


### 3 Advanced Part 

```{r}
interval = 24 * 60
minutes = seq(1, interval)

# -1 so we can have an actual value for the last predicted 10-min forward returns
days = (nrow(prices) / interval) - 1

H <- c(10, 30)
I <- c(1, 2, 3)

# H-min * I assets for forward returns, backward returns, and volatility
# + 1 * interval for asset 1 prices t = [1, 1440], and + 1 for actual value
# of 10-minute forward returns for the last minute of the day
day_prices = matrix(nrow=days, ncol=((length(I)*(length(H)) + 2)*interval))
```

```{r}
for (d in 1:days) {
  col_num = 2
  idxs = (d * interval) + minutes
  
  end = d * interval
  f_ret = prices[end + 10, 2]
  day_prices[d, 1] = round((f_ret - prices[end, 2])/ prices[end, 2], 4) 
  
  day_prices[d, col_num:(col_num + interval - 1)] = round(prices[idxs, 2], 4)
  col_num = col_num + interval
  
  for (i in I) {
    asset_prices = prices[idxs, i+1]
    
    for (h in H) {
      prev_returns = asset_prices[pmax(minutes - h, 1)]
      returns = round((asset_prices - prev_returns) / prev_returns, 4)
      
      day_prices[d, col_num:(col_num + interval - 1)] = returns
      col_num = col_num + interval
    }
    
    #for (h in H) {
    #  forward_returns = asset_prices[pmin(minutes + h, interval)]
    #  returns = round((forward_returns - asset_prices) / asset_prices, 4)
      
    #  day_prices[d, col_num:(col_num + interval - 1)] = returns
    #  col_num = col_num + interval
    #}
    
    # for (h in H) {
    if (i == 1) {
      for (j in 2:interval) {
        day_prices[d, col_num] = round(sd(asset_prices[max(1, j - 9):j]), 4)
        col_num = col_num + 1
      }
    }
  }
}
```

```{r}
col_names = c("Asset_1_10-FRet_End")
col_names = append(col_names, paste("Asset_1_Price_t", minutes, sep=""))

for (i in I) {
  for (h in H) {
    col_names= append(col_names, paste("Asset_",i,"_",h,"-BRet_t", minutes, sep=""))
  }
  
  #for (h in H) {
  #  col_names = append(col_names, paste("Asset_",i,"_",h,"-FRet_t", minutes, sep=""))
  #}
  
  # Exclude NA columns
  if (i == 1) {
    col_names = append(col_names, paste("Asset_",i,"_10-Vol_t", minutes[-1], sep=""))
  }
}
```

```{r}
DayPrices = data.frame(day_prices)
colnames(DayPrices) <- col_names
# write.csv(DayPrices, 'day_prices.csv', row.names = FALSE)
```

```{r}
set.seed(14532)
train_proportion = 0.7

# Select first 70% of data as training data
train_size <- floor(nrow(DayPrices) * train_proportion)
train_ids = 1:train_size

prices_train <- DayPrices[train_ids, ]
prices_test <- DayPrices[-train_ids, ]

# options("expressions"=5e5)
# mem = memory.limit(size=16000000)

train_mat = model.matrix(`Asset_1_10-FRet_End` ~ ., data=prices_train)
test_mat = model.matrix(`Asset_1_10-FRet_End` ~ ., data=prices_test)
```


```{r}
library(glmnet)

cv.ridge = cv.glmnet(train_mat, prices_train$`Asset_1_10-FRet_End`, alpha=0, nfolds=10)
ridge.mod = glmnet(train_mat, prices_train$`Asset_1_10-FRet_End`, alpha=0)
```

```{r}
bestlam_ridge = cv.ridge$lambda.min

ridge.pred_train = predict(cv.ridge, s=bestlam_ridge, newx=train_mat)
ridge.pred_test = predict(cv.ridge, s=bestlam_ridge, newx=test_mat)

in_sample_adv_ridge = cor(ridge.pred_train, prices_train$`Asset_1_10-FRet_End`)
out_sample_adv_ridge = cor(ridge.pred_test, prices_test$`Asset_1_10-FRet_End`)
  
cat("In-sample correlation for advanced ridge:", in_sample_adv_ridge,
  "\nOut-sample correlation for advanced ridge:", out_sample_adv_ridge)
```


```{r}
cv.lasso = cv.glmnet(train_mat, prices_train$`Asset_1_10-FRet_End`, alpha=1, nfolds=10)
lasso.mod = glmnet(train_mat, prices_train$`Asset_1_10-FRet_End`, alpha=1)
```

```{r}
bestlam_lasso = cv.lasso$lambda.min

### returns nonzero coefs
rownames(coef(cv.lasso, s = 'lambda.min'))[coef(cv.lasso, s = 'lambda.min')[,1]!= 0]

lasso.pred_train = predict(cv.lasso, s=bestlam_lasso, newx=train_mat)
lasso.pred_test = predict(cv.lasso, s=bestlam_lasso, newx=test_mat)

in_sample_adv_lasso = cor(lasso.pred_train, prices_train$`Asset_1_10-FRet_End`)
out_sample_adv_lasso = cor(lasso.pred_test, prices_test$`Asset_1_10-FRet_End`)
  
cat("In-sample correlation for advanced ridge:", in_sample_adv_lasso,
    "\nOut-sample correlation for advanced ridge:", out_sample_adv_lasso)
```

```{r}
library(pls)

cv.pcr <- pcr(`Asset_1_10-FRet_End` ~ ., data=DayPrices, 
              subset = train_ids, validation = "CV")
validationplot(cv.pcr, val.type = "MSEP", legendpos = "topright")
```

```{r}
cverr <- RMSEP(cv.pcr)$val[1,,]
ncomp_min <- unname(which.min(cverr) - 1)

pcr.pred_train = predict(cv.pcr, DayPrices[train_ids, ], ncomp = ncomp_min)
pcr.pred_test = predict(cv.pcr, DayPrices[-train_ids, ], ncomp = ncomp_min)

in_sample_adv_pcr = cor(pcr.pred_train, prices_train$`Asset_1_10-FRet_End`)
out_sample_adv_pcr = cor(pcr.pred_test, prices_test$`Asset_1_10-FRet_End`)
  
cat("In-sample correlation for advanced PCR:", in_sample_adv_pcr,
  "\nOut-sample correlation for advanced PCR:", out_sample_adv_pcr)
```



```{r}
new_prices = matrix(nrow=1, ncol=((length(I)*(length(H)) + 2)*interval))

prediction <- function(asset_prices) {
  col_num = 2
  new_prices[1, col_num:(col_num + interval - 1)] = round(asset_prices[minutes, 2], 4)
  col_num = col_num + interval
  
  for (i in I) {
    for (h in H) {
      prev_returns = asset_prices[pmax(minutes - h, 1), i]
      returns = round((asset_prices[, i] - prev_returns) / prev_returns, 4)
      
      new_prices[1, col_num:(col_num + interval - 1)] = returns
      col_num = col_num + interval
    }
    
    if (i == 1) {
      for (j in 2:interval) {
        new_prices[1, col_num] = round(sd(asset_prices[max(1, j - 9):j, i]), 4)
        col_num  = col_num + 1
      }
    }
  }
  
  return(unname(predict(cv.lasso, s=bestlam_lasso, newx=new_prices)[1,]))
}
```

```{r}
vals = c()

options(digits.secs=6)
start_time = Sys.time()

for (i in 1:10000) {
  vals = c(vals, prediction(prices[1:1440, 2:4]))
}

end_time = Sys.time()
```

```{r}
temp = prices[, 2]
prices[, 2] = prices[, 4]
prices[, 4] = temp
```

```{r}
interval = 24 * 60
minutes = seq(1, interval)

# -1 so we can have an actual value for the last predicted 10-min forward returns
days = (nrow(prices) / interval) - 1

H <- c(10, 30)
I <- c(1, 2, 3)

predicted_returns = c()
expected_returns = c()

for (d in 1:days) {
  idxs = (d * interval) + minutes
  
  # Expected value
  end = d * interval
  f_ret = prices[end + 10, 2]
  expected_returns[d] = round((f_ret - prices[end, 2])/ prices[end, 2], 4) 
  
  # Predicted value
  predicted_returns[d] = prediction(prices[idxs, 2:4])
}

pred_cor = cor(expected_returns, predicted_returns)
```
